---
title: "Assignment_3"
author: "Priya Shaji"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##R Character Manipulation and Date Processing {.tabset}

### Question 3
3. Copy the introductory example. The vector name stores the extracted names.

(a)

```{r}
library(stringr)
library(knitr)
```

Introductory Example

```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```

Extract vectors, Vector name stores the extracted name
```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```

Replace the title and the middle name for name dataframe with blank

```{r}
remove_title_middle <- str_replace(name,"([[:alpha:]]{1,3}\\.\\s)", "")
```

Rearrange the vector so that all elements conform to the standard first_name last_name.

Reverse the first and last name where necessary

```{r}
firstName_lastName <- str_replace(remove_title_middle, "([[:alpha:]]+), ([[:alpha:]]+)", "\\2 \\1")

kable(list(data.frame(name, firstName_lastName)), caption = "Reorder \"name\" so it conform to the standard first and last name format.")
```

As we can see, all the names are separated as the standard format, firstName, lastName

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

```{r}
title_char <- str_detect(name, "[[:alpha:]]{2,3}\\. ")
kable(list(data.frame(name, title_char)), caption = "Detect if names in \"name\" has a title.")
```

    

(c) Construct a logical vector indicating whether a character has a second name.

```{r}
second_name <- str_detect(name, " [[:alpha:],]{1,}")
kable(list(data.frame(name, second_name)), caption = "Detect if names in \"name\" has a second name.")
```

All the names in the name vector has a second name, therefore, the resulting table has TRUE for all the inputs.

### Question 4
4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.
(a) [0-9]+\\$

This regular expression is matched by a string which starts with a number/s [0-9], + sign means that the preceeding item will be matched one or more time, \\$ ends with a $ sign

example:
```{r}
str_1<- c("123$", "abc$", "-1098$", "test$", 27)
test_1<- unlist(str_extract_all(str_1, "[0-9]+\\$"))
test_1
```


To detect if the above regular expression works:

```{r}
test_1 <- str_detect(str_1, "[0-9]+\\$")
test_1
```



(b) \\b[a-z]{1,4}\\b

This regular expression is matched by a string which is followed by lowe case alphabets which ranges between a to z.We are asking the function for all instances where this sequence appears at least once, but at most four times. \\b indicates the word edges
```{r}
str_2<- c("aa", "bzdg", "bnewyork", "jeny")
test_2<- unlist(str_extract_all(str_2, "\\b[a-z]{1,4}\\b"))
test_2
```

To detect if the above regular expression works:

```{r}
test_2 <- str_detect(str_2, "\\b[a-z]{1,4}\\b")
test_2
```

(c) .*?\\.txt$

This regular expression is matched by any string which ends by .txt

```{r}
str_3<- c("aa.txt", "bzdg", "$bnewyork", "sneha.txt")
test_3<- unlist(str_extract_all(str_3, ".*?\\.txt$"))
test_3
```

To detect if the above regular expression works:

```{r}
test_3 <- str_detect(str_3, "\\b[a-z]{1,4}\\b")
test_3
```


(d) \\d{2}/\\d{2}/\\d{4}

This regular expression is matched by a number pattern dd/dd/dddd

```{r}
str_4<- c("12/12/2019", "11-22/2018", "11/22/2018", "sneha.txt")
test_4<- unlist(str_extract_all(str_4, "\\d{2}/\\d{2}/\\d{4}"))
test_4
```

To detect if the above regular expression works:
```{r}
test_4 <- str_detect(str_4, "\\b[a-z]{1,4}\\b")
test_4
```


(e) <(.+?)>.+?</\\1>

This regular expression is matched by a pattern with open and closed tags, like the html tags.
```{r}
str_5 <- c("<b> qwerty </b>", "<h1>priya <h1>", "<>test</tag>", "<tag> helloworld </tag>")
test_5 <- unlist(str_extract_all(str_5, "<(.+?)>.+?</\\1>"))
test_5
```


To detect if the above regular expression works:
```{r}
test_5 <- str_detect(str_5, "<(.+?)>.+?</\\1>")
test_5
```


### Question 9
9. The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.
     clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr
     
```{r}
str_test <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
str_extract_all(str_test, "[a-z]")
```
     
```{r}

str_extract_all(str_test, "[A-Z]")

```

     
The hidden message is revealed.
"CONGRATULATIONS YOU ARE A SUPER NERD"
     