---
title: "Assignment1_Mushroom"
author: "Priya Shaji"
date: "January 31, 2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read the file in table format and create a data frame from it
```{r}
Mushroom_Family <- read.csv(url("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"), header = FALSE)

```
## Mushroom_Family is the dataframe created. Dimensions of the dataframe:
```{r}
dim(Mushroom_Family)
```
## The first six rows of the data frame are:
```{r}
head.matrix(x = Mushroom_Family, 6L)

```
## Mushroom_Sibling is the subset of the Mushroom_Family dataframe from the selected columns:
```{r}
Mushroom_Sibling <- Mushroom_Family[,c(1,4,6,10,11,23)]

```
## Renaming the column names of the subset dataframe:
```{r}
names(Mushroom_Sibling) = c("Type", "Cap_Color", "Odor", "Gill_Color", "Stem_Shape","Habitat")

```
## Column names of Mushroom_Sibling dataframe are renamed:
```{r}
names(Mushroom_Sibling)

```

## Replace the abbreviation used in column Type:
1. Convert Columns to characters
```{r}
Mushroom_Sibling$Type <- as.character(Mushroom_Sibling$Type)

```
2. Replace abbreviation with meaningful names
```{r}
Mushroom_Sibling$Type[Mushroom_Sibling$Type == "e"] <- "edible"
Mushroom_Sibling$Type[Mushroom_Sibling$Type == "p"] <- "poisonous"

```
3.If needed to maintain columns as factors 
```{r}
Mushroom_Sibling$Type <- as.factor(Mushroom_Sibling$Type)

```
##Display first six rows of the dataframe to verify the change
```{r}
head.matrix(Mushroom_Sibling, 6L)

```
##Replace abbreviations of colum Cap_Color 
```{r}
Mushroom_Sibling$Cap_Color <- as.character(Mushroom_Sibling$Cap_Color)
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "n"] <- "brown"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "b"] <- "buff"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "c"] <- "cinnamon"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "g"] <- "gray"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "r"] <- "green"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "p"] <- "pink"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "u"] <- "purple"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "e"] <- "red"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "w"] <- "white"
Mushroom_Sibling$Cap_Color[Mushroom_Sibling$Cap_Color == "y"] <- "yellow"
Mushroom_Sibling$Cap_Color <- as.factor(Mushroom_Sibling$Cap_Color)

```
## Replace Abbreviations of column Odor
```{r}
Mushroom_Sibling$Odor <- as.character(Mushroom_Sibling$Odor)
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "a"] <- "almond"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "l"] <- "anise"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "c"] <- "creosote"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "y"] <- "fishy"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "f"] <- "foul"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "m"] <- "musty"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "n"] <- "none"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "p"] <- "pungent"
Mushroom_Sibling$Odor[Mushroom_Sibling$Odor == "s"] <- "spicy"
Mushroom_Sibling$Odor <- as.factor(Mushroom_Sibling$Odor)

```

## Replace abbreviation of column Gill_Color
```{r}
head.matrix(Mushroom_Sibling, 3L)
Mushroom_Sibling$Gill_Color <- as.character(Mushroom_Sibling$Gill_Color)
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "k"] <- "black"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "n"] <- "brown"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "b"] <- "buff"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "h"] <- "chocolate"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "g"] <- "gray"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "r"] <- "green"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "o"] <- "orange"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "p"] <- "pink"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "u"] <- "purple"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "e"] <- "red"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "w"] <- "white"
Mushroom_Sibling$Gill_Color[Mushroom_Sibling$Gill_Color == "y"] <- "yellow"
Mushroom_Sibling$Gill_Color <- as.factor(Mushroom_Sibling$Gill_Color)

```
## Replace abbreviation of column Stem_Shape
```{r}
Mushroom_Sibling$Stem_Shape <- as.character(Mushroom_Sibling$Stem_Shape)
Mushroom_Sibling$Stem_Shape[Mushroom_Sibling$Stem_Shape == "e"] <- "enlarging"
Mushroom_Sibling$Stem_Shape[Mushroom_Sibling$Stem_Shape == "t"] <- "tapering"
Mushroom_Sibling$Stem_Shape <- as.factor(Mushroom_Sibling$Stem_Shape)

```
## Replace abbreviations of column Habitat
```{r}
Mushroom_Sibling$Habitat <- as.character(Mushroom_Sibling$Habitat)
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "g"] <- "grasses"
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "l"] <- "leaves"
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "m"] <- "meadows"
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "p"] <- "paths"
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "u"] <- "urban"
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "w"] <- "waste"
Mushroom_Sibling$Habitat[Mushroom_Sibling$Habitat == "d"] <- "woods"
Mushroom_Sibling$Habitat <- as.factor(Mushroom_Sibling$Habitat)

```
##Display the table to verify:
```{r}
head.matrix(x = Mushroom_Sibling, 6L)

```
##By exploring and analyzing the dataset and the data description of Mushrooms ,the following conclusions can be drawn:
1) Edible and poisonous mushrooms can be analyzed based on two strong factors:
Odor and Habitat
With Odor being the strongest factor.

2) Therefore Let's calculate the frequencies of both the factors affecting the edibility of the mushrooms:
```{r}
table = table(Mushroom_Sibling$Type, Mushroom_Sibling$Odor,Mushroom_Sibling$Habitat)
result_data = as.data.frame(table)
names(result_data)[1] = 'Type'
names(result_data)[2] = 'Odor'
names(result_data)[3] = 'Habitat'
result_data
```


