---
title: "Assignment_5"
author: "Priya Shaji"
date: "2/28/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tidying and Transforming Data {.tabset}
### loading the required packages
```{r}
#install.packages("tidyr")
#install.packages("dplyr")
library(tidyr)
library(dplyr)
```

### Dataset

#### Create a dataframe and load the dataset from the .csv file

```{r}
df_flight_wide <- read.csv("/Users/priyashaji/Documents/cuny msds/Spring'19/data 607/assignments/assignment_5/FlightsInfo - Sheet1.csv",header=TRUE,stringsAsFactors = FALSE)
df_flight_wide
```

### Tidying and Transforming data 

Using functions provided by tidyr and dplyr packages
The current dataframe: df_flight is a wide format of data
We will use gather() function of tidyr package to make the data in long format


gather() function uses the folowing arguments:

 data: Data object
 
 key: Name of new key column (made from names of data columns)
 
 value: Name of new value column
 
 Names of source columns that contain values
 
 factor_key: Treat the new key column as a factor (instead of character   vector)

```{r}
df_flight_long <- gather(df_flight_wide, city, number, Los.Angeles:Seattle, factor_key=TRUE)
df_flight_long
```

rename the columns

```{r}
colnames(df_flight_long)[2] <- "Status" 
colnames(df_flight_long)[1] <- "Airline"
df_flight_long$Airline <- factor(df_flight_long$Airline)
df_flight_long
```


As we see in the above dataframe, "Airline" has alternate missing values. In order to fill the missing values from the Airline name previous to it, we use na.locf(.) function. Therefore, we first install the package "zoo"


na.locf(.) :

Last Observation Carried Forward
Generic function for replacing each NA with the most recent non-NA prior to it.




```{r}

#install.packages("zoo")
library(zoo)

df_flight_long[df_flight_long$Airline=="","Airline"] <- NA
df_flight_long <- df_flight_long %>%
  do(na.locf(.))
df_flight_long

# convert the column "number" from type "character" to type "numeric"
df_flight_long$number<-suppressWarnings(as.numeric(df_flight_long$number))
is.numeric(df_flight_long$number)
```



Our dataframe is in long format now, we can apply dataset

We will use the spread() function.


spread() returns a copy of your data set that has had the key and value columns removed. In their place, spread() adds a new column for each unique value of the key column. These unique values will form the column names of the new columns. spread() distributes the cells of the former value column across the cells of the new columns and truncates any non-key, non-value columns in a way that prevents duplication.

```{r}
df_flight <- spread(df_flight_long, "Status", "number") 

is.factor(df_flight$Airline)

is.numeric(df_flight$`on time`)
is.numeric(df_flight$delayed)

is.factor(df_flight$city)

df_flight


```


### Analysis of flight status

Analysis to compare the arrival delays for the two airlines


Using the pipe(%>%) operator to do the analysis

```{r}
delay_analysis<- as.data.frame(df_flight %>%
  mutate(total = delayed + `on time`) %>%
  group_by(Airline,city) %>%
  summarise(delayed_total = sum(delayed), 
            on_time_total = sum(`on time`),
            count_total = sum(total),
            percent_ontime = round(on_time_total/count_total*100,digit = 2),
 percent_delayed = round(delayed_total/count_total*100,digit = 2)         ))
delay_analysis
```



```{r}
library(ggplot2)
delays <- ggplot(delay_analysis,aes(x = city, y = percent_delayed))
suppressWarnings(delays + geom_bar(stat = "identity", position="dodge", aes(fill = Airline)))
```

#### Conclusion

As we can see by the graph, the delayed rate of AM WEST is higher than ALASKA Airline. 

